\maketitle
%\makedeclaration

\begin{abstract} % 300 word limit
\normalsize
Nowadays the volume of information available on the internet have been increasing rapidly with the explosive growth of the World Wide Web, e-Commerce and other online services. Many websites usually offer recommendation systems to help customers to find items that satisfy their needs. However, such systems usually act in a greedy manner by recommending items with the highest user ratings and without actively gathering information on current user preferences.

A successful recommender system must balance the need to explore user preferences and to exploit this information for recommendation. Reinforcement learning has demonstrated to be a potentially effective approach in the domain by generating recommendations as a sequential optimization problem, nevertheless, it has received relatively little attention and found only limited application as its computational complexity made the problem intractable when there is a high number of items to recommend.

The main objective of this project is to demonstrate that recent advances in deep neural networks and reinforcement learning models are useful to build a model that performs well under a recommendation system environment and are able generate good recommendations to users over a large set of items. After defining the recommendation task under the Open AI gym framework\footnote{\url{https://gym.openai.com/}}, the performance of reinforcement learning agents based on the Deep Deterministic Policy Gradient algorithm with k-nearest neighbors and Factorization Machine variants are compared. [***](Missing experiments and conclusions)

\end{abstract}

\begin{acknowledgements}

Firstly, I would like to express my sincere gratitude to my supervisor and programme director Dr. Jun Wang for his advice and comments on this work, and for being our mentor at University College London.

I would also like to thank to the Secretary of Higher Education, Science, Technology and Innovation (SENESCYT), and especially, to Dr. Rafael Correa Delgado, president of the Republic of Ecuador, for believing in the skills of all Ecuadorian citizens studying in the best Universities around the world. Your trust and our sacrifice will be transformed into innovation that will turn our Country into a republic of opportunities. 

Finally, I would like to thanks to my mom, friends and family for their support and encouragement throughout my studies.



\end{acknowledgements}


\setcounter{tocdepth}{2} 
% Setting this higher means you get contents entries for
%  more minor section headers.

\tableofcontents
\listoffigures
\listoftables

